Mitigating Interest Flooding DDoS Attacks in NDN

### 2013-01-11 03-interest-flooding.tex
Edits to Section III
git-svn-id: https://ndnpub.parc.com/ndn-papers/NDN-SEC/dos-protection@897 df1a90cd-247b-4364-bdaf-287a704eb499

\paragraph{What is it?}

Unlike data packets, Interest packets in NDN are unsolicited, consume state in intervening routers and are routed based on content name prefixes in the network. Such properties makes Interest packets an effective tool in launching DoS attacks in NDN. An attacker or a set of distributed attackers can inject excessive amounts of Interests to overload the network and cause service disruptions for legitimate users (Fig.~\ref{fig:flooding example}). We coin in the term \emph{Interest flooding} to refer to these network-level flooding attacks.

@@

\paragraph{How it works?}

The Interest flooding attacks can distrupt service quality in NDN network in two ways: it \emph{creates network congestion} and \emph{exhaust resources on routers}.

Like any packet in a network, Interest packets consume a portion of network capacity. Thus, Interest Flooding attacks might cause congestion and the drop of legitimate packets in the network. Although congestions might occur anywhere in the network, a coordinated DDoS attack would likely target one specific namespace to concentrate attack traffic in certain segments of the network. That is feasible for an attacker as the name prefixes loosely correspond to a network location or sets of network locations in network.

As NDN routers maintain per-packet states for each forwarded Interest (i.e., an entry in its Pending Interests Table), an excessive volume of malicious Interests can lead to exhaustion of routers' memory. When a router has insufficient memory to create a new PIT entry, it has to drop new incoming Interests.
% there is no point in forwarding the Interest further.
% Otherwise, if Data comes back, it would be considered unsolicited and dropped anyways.
Therefore, the service for legitimate users can be disrupted due to limited memory resources in routers without saturating communication links to their limits.

@@

Because communication in NDN is content-centric, it is unlikely for an adversary to be able to target specific routers or end-hosts, as neither of them is required to have a globally routable network identity/address. However, an adversary can target a specific namespace.
For example on Fig.~\ref{fig:flooding example}, if the Data producer is the exclusive owner of \ndnName{/foo/bar} namespace and is single-homed to router B, both router B and the Data producer would receive all Interests for \ndnName{/foo/bar/\ldots} that cannot be otherwise satisfied from in-network caches. In the case of multi-homed users with rich connectivity through multiple providers, NDN would likely to cope better with DDoS attacks due its native multi-path and adaptive forwarding~\cite{adaptive-forwarding}.

@@

To efficiently implement an Interest flooding attack targeting a namespace (e.g., /newtorktimes/), an adversary needs to make sure that (1) the expressed Interests are getting as close to the Data producer as possible, and (2) the corresponding PIT entries (state) are kept on NDN routers for as long as possible.\footnote{While NDN allows users to specify lifetime of the Interest inside the Interest packet~\cite{ndn-conext,ndn-tr}, it is router's decision for how long it is willing to keep the PIT entry.  For example, and we assume this in the rest of the paper, the maximum time for which any Interest is admitted is one second.}

In order to achieve both requirements most effectively, malicious Interests should avoid collapsing (i.e, request a content that is different than existing PIT entries in the network) and cache hits (i.e., request unpopular or non-existing content). In other words, to maximize effect of an DDoS attack, each malicious Interest generated by an individual attack bot should request a non-existing content with a unique name (\emph{unique junk Interests}). In this paper, we exclusively focus on this particular attack strategy as it maximizes the damaging effect of each malicious Interest in the network and the same or similar countermeasures apply to all less-effective strains of the attack.

In the rest of this paper, we use the general term \emph{Interest flooding attack} to refer to the above described attack and assume an attacker that is limited to controlling a botnet of end-hosts only (i.e, we assume the routers in the network and the computers in the victim domain are not compromised).

>>>>

%\paragraph{What is it?}

Unlike Data packets, Interest packets in NDN are unsolicited, consume state in intermediate routers and are routed through the network based on their content name prefix. These properties make Interest packets an effective tool in launching DoS and DDoS attacks in NDN. An attacker or a set of distributed attackers can inject excessive amounts of Interests to overload the network and cause service disruptions for legitimate users (Fig.~\ref{fig:flooding example}). We coin the term \emph{Interest flooding} to refer to these network-level flooding attacks.

@@

%\paragraph{How does it work?}

Since communication in NDN is content-centric, it is unlikely that an adversary can target specific routers or end-hosts, as neither is required to have a globally routable network identity/address. However, an adversary can target a specific namespace.
For example in Fig.~\ref{fig:flooding example}, if the Data producer is the exclusive owner of \ndnName{/foo/bar} namespace, both router B and the Data producer would receive all Interests for \ndnName{/foo/bar/\ldots} that cannot be otherwise satisfied from in-network caches. \footnote {In our example, Data producer is single-homed. In the case of multi-homed producers with rich connectivity through multiple providers, NDN would likely to cope better with DDoS attacks due its native multi-path and adaptive forwarding~\cite{adaptive-forwarding}.} These large volume of Interests can distrupt service quality in NDN network in two ways: they \emph{create network congestion} and \emph{exhaust resources on routers}.

Similar to packets in traditional networks, Interest packets in NDN consume a portion of network capacity. A large number of Interest packets might cause congestion and lead to legitimate packets being dropped in the network. Although congestion might occur anywhere in the network, a coordinated DDoS attack would likely target one specific namespace to concentrate attack traffic in certain segments of the network, typically closest to the publisher serving that namespace since name prefixes in NDN loosely correspond to a network location or sets of network locations.

As NDN routers maintain per-packet states for each forwarded Interest (i.e., an entry in its PIT), an excessive amount of malicious Interests can lead to exhaustion of a routers' memory. When a router has insufficient memory to create a new PIT entry, it is forced to drop new incoming Interests, thus disrupting service for legitimate users.
% there is no point in forwarding the Interest further.
% Otherwise, if Data comes back, it would be considered unsolicited and dropped anyways.
%Therefore, service for legitimate users can be disrupted due to limited memory resources in routers without saturating communication links to their limits.

@@

@@

To efficiently implement an Interest flooding attack targeting a specific namespace (e.g., /newyorktimes/), an adversary needs to make sure that (1) the expressed Interests are routed as close to the Data producer as possible, and (2) the corresponding PIT entries (state) are stored at intermediate NDN routers for as long as possible.\footnote{While NDN allows users to specify lifetime of the Interest inside the Interest packet~\cite{ndn-conext,ndn-tr}, it is router's decision for how long it is willing to keep the PIT entry.  For example, and we assume this in the rest of the paper, the maximum time for which any Interest is admitted is one second.}
In order to achieve both requirements effectively, malicious Interests should avoid collapsing (i.e, request a content that is different than existing PIT entries in the network) and avoid being served from caches of intermediate routers (i.e., request unpopular or non-existing content). In other words, to maximize effect of an DDoS attack, each malicious Interest generated by an individual attacker bot should request a non-existing content with a unique name (\emph{unique junk Interests}). In this paper, we exclusively focus on this particular attack strategy as it maximizes the damage from each malicious Interest in the network. Other less effective strains of this attack can be mitigated by applying the same or similar countermeasures.

In the rest of this paper, we use the general term \emph{Interest flooding attack} to refer to the above described attack and assume an attacker that is limited to controlling a botnet of end-hosts only i.e., we assume the routers in the network and the computers in the victim domain are not compromised.
